\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-20 europa-basics.tex
added EUROPA constraints in the form of a table

@@

{\footnotesize
  \begin{quote}
[Planning] \emph{is an abstract, explicit deliberation process that chooses and
organizes actions by anticipating their expected outcomes. This
deliberation aims at achieving as best as possible some prestated
objectives. Automated planning is an area of Artificial Intelligence
(AI) that studies this deliberation process computationally.} -- from
\textbf{Automated Planning Theory and Practice} by Ghallab, Nau and
Traverso \cite{ghallab04} 
\end{quote}
}

@@

For such a robot the milk buying plan at the store might look like
\comment{Figure needed}.

% \eu is a general purpose AI planning toolkit developed at NASA Ames.  
% TODO: Talk about NASA missions where \eu has been used.

@@

\paragraph \textit{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \comment{citations needed}, a project
consisting of a set of activities must be scheduled in a way that
satisfies minimum and/or maximum temporal separation constraints. The
activity schedule must also respect fixed limits on the availability of
resources required to perform each activity. In addition to satisfying
temporal and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \comment{citation} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to be arc consistent if for every
value in $x$'s domain, there is a value in $y$'s domain that is
consistent with all the constraints that connect $x$ and $y$.  A naive
approach to ensuring arc consistency consists of cycling through all
the variable pairs and performing constraint propagation until there
are no variable domain changes. The AC-3 algorithm \cite{mackworth77},
a popular implementation, improves over the naive approach by ignoring
constraints whose variables were not affected since the last
iteration.

Arc Consistency and other, more stringent consistency definitions can be
used to efficiently prove that a CP problem is unsatisfiable. However,
they generally cannot prove that a CP problem is satisfiable and
provide specific variable assignments that constitute solutions. For
that, search algorithms like global backtracking \comment{citation
  needed} or local search are used; these use consistency as an
efficient way to prune the search space \comment {citation CP
  Handbook}. In theory, solving a CP problem is NP-Hard
\cite{ghallab04}, but often very efficient in practice using a number
of consistency and search algorithms that are well understood.

CP is usually implemented as part of a programming language and
constraints are usually represented as objects \cite{puget95}. Any
constraint that the user is able to reason about can be introduced into the system and, as long as the constraint
propagation protocols specified by the host CP system are enforced, it
will be indistinguishable from any other "primitive'' CP constraint,
such as $\leq$ or $\geq$.

\subsection{Constraint-Based Attribute and Interval Planning}
\label{sec:europa:cp}

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
mutual exclusion and concurrency using propositions \comment{citation
  needed}. While CSP representations have traditionally had an edge in
formulating and representing planning problems, 
% It is straightforward to represent and reason about all of
% those elements using a CSP representation, as a result there has been
% some work on doing automated planning while taking advantage of CSP
% (TODO: ref). Traditionally, the entire planning problem is translated
% into a CSP and then solved using traditional CSP methods, this
leading to a formulation where action choices and relationships are
represented as variables and/or constraints, there are at least two
major drawbacks:

@@

Constraint-based Planning and specifically Constraint-Based Attribute
and Interval planning (CAIP) \cite{mus94,frank2003} is intended to
close that gap. On one hand, it takes advantage of a CSP
representation; on the other it uses attributes and intervals to
maintain an explicit representation of the elements relevant to
planning. This makes it easier to write algorithms that search for an
reason abut plans. The main elements of CAIP are:

\begin{enumerate}
	\item Intervals: \comment{needs to be defined}
	\item Attributes:
	\item Domain Constraints an Configuration rules:
\end{enumerate}

\comment{Explain how planning problem instances and their solutions are represented}

\subsection{\eu Plan Representation}
\label{sec:europa:pr}

\eu is an implementation of the CAIP framework.

\textit{Variables} Values that need to be represented to describe the
problem domain and over which we may want to specify constraints.

\textit{Objects} The things we wish to describe and refer to in a
domain are considered Objects. As in the case with object-oriented
analysis and design, one can seek out the nouns in any domain
description to find likely objects. In a Mars rover example, we might
consider the satellite, the camera, and the attitude controller (which
orients the robot for taking pictures) to be objects. Objects have
state and behavior. For example, a camera can be:

    off,
    ready, or
    taking a picture.

An attitude controller can be:

    pointing at a position, or
    turning from one position to another.

    An object is an instance of a class. In \eu we model using the
    abstraction of a class to speak about all instances having certain
    properties of state and behavior. In order to describe such state
    and behavior we build on the formalism of first order logic.

\textit{Tokens}

A predicate defines a relation between objects and properties. In \eu,
we define such relations between variables whose domains are sets of
objects and sets of properties. For example, we might use a predicate
\texttt{Pointing($a,p$)} to indicate that a rover's attitude controller
$a$ is pointing at position $p$. Note that $a$ is a variable which may
have a number of possible values in a problem with multiple
satellites. Similarly, $p$ is a variable whose values are the set of
possible positions. In a grounded plan, single values will be
specified for each variable as we saw in the case of a solved
Constraint Satisfaction Problem.

In general, it is not sufficient to state that a predicate is true
without giving it some temporal extent over which it holds. Predicates
that are always true can be thought to hold from the beginning to the
end of time. However, in practice, the temporal extent of interest
must be defined with timepoints to represent its start and end. So we
might want to write \texttt{Pointing($a,s,e,p$)} to indicate that the
attitude controller a is pointing at position $p$ from time $s$ to
time $e$. In fact, this pattern of using such predicates to describe
both state and behavior of objects is so prevalent in \eu that we have
introduced a special construct called a Token which has the built in
variables to indicate the object to which the statement principally
applies and the timepoints over which it holds. In \eu, all predicate
instances are Tokens.

\comment{talk about actions}

A Token is an instance of a predicate or an action and is defined over
a temporal extent. Every token has five built-in variables:

start: The beginning of the temporal extent over which the predicate
is defined.  end: The end of the temporal extent over which the
predicate is defined.  duration: The constraints start + duration =
end is enforced automatically.  object: made. The set of objects to
which a token might apply. In a grounded plan each Token applies to a
specific Object, reflecting the intuition that we are using Tokens to
describe some aspect of an Object (i.e. its state or behavior) in
time. However, in a partial plan, the commitment to a specific object
may not yet have been made.

state: Tokens can be ACTIVE, INACTIVE, MERGED, or REJECTED. The state
variable captures the token's current state and its reachable states
through further restriction. See the Token State Model discussion
below for details.

\textit{Objects (continued)}

As we have noted, Tokens describe some aspect of an Object in
time. Objects thus may have many Tokens in a plan in order to describe
their state and behavior throughout all points in time of the
plan. Within this general framework, we note a few particulars:

Static Facts: Classes without predicates lead to Objects without
Tokens. This arises where an objects state or behavior is independent
of time. For example, a domain may have a set of locations and/or
paths for which there is nothing more to say than that they exist.

Timelines: Often objects in a domain must be described by exactly one
token for every given timepoint in the plan. Such objects are so
common that we provide a special construct in \eu to extend these
semantics to derived classes. Any instances of a class derived from a
Timeline will induce ordering requirements among its tokens in order
to ensure no temporal overlap may occur among them. See below for
further discussion.

Resources: Metric resources, e.g. the energy of a battery or the
capacity of a cargo hold, are objects with an explicit quantitative
state in time and with a circumscribed range of changes that can occur
to impact that state i.e. produce, consume, use, change. These changes
are captured as tokens. Resources are such a common requirement for
\eu users that special constructs are also provided for
them. Instances of classes derived from a Resource will induce
ordering requirements on their Tokens in order to ensure that the
level of the resource remains within specified limits.

Timelines

It may be sufficient to maintain only a partial-order among
tokens. However, it is often the case that tokens represent states and
actions of a single object in the system. Such tokens are typically
mutually-exclusive. \eu uses a Timeline structure developed in (??)
to concisely capture system components whose behavior is described
over time in this manner.

\comment{This is the first instance that this domain is being
  introduced. Either we need more explanation or it ought to be
  removed.} For example, consider the tire-world domain. The tire was
located in the trunk with the predicate tireLocated(Trunk) and located
on the ground with the predicate tireLocated(Ground). Clearly, the
same tire cannot be in both places at once. A Timeline provides a
simple method for aggregating the statements about a tire such that
they are mutually exclusive.

Figure 1: A Timeline for a Tire

Figure 1 illustrates how a Timeline can be used to describe the
whereabouts of a tire. The tire is an object in the tire-world
represented as a timeline. It has predicates associated with it which
can describe its states and actions over time. The predicate names
need not contain the tire prefix; this is now implicit since the
tokens are assigned to a specific tire instance (i.e. an instance of a
Timeline). While Timelines are a useful element of the \eu planning
paradigm, they are not essential. A more general notion of a system
Object can be used which does not impose restrictions of
mutual-exclusion or non-zero duration. This can be important in
supporting partial-order planning.

For this example, the state of the tire is specified using 3
contiguous tokens. The end and start time-points are related by an
equality constraint. A Moving predicate has been introduced to cover
the transition from one location to another. It takes 2 location
arguments. Notice that precedence relationships exist between tokens
such that they cannot overlap but the start and end times may remain
flexible. To illustrate this, sample times are included. Assume a
total time range of interest between 0 and 1000. In addition, tokens
on Timelines have a minimum duration of 1. As a result the values
shown are the most flexible possible values for the start and end of
each token. There are a number of advantages of allowing this
flexibility. First, the basic structure of the plan can be developed
without over-committing to specific times. If it is not necessary for
the tire to be on the ground at time-step 4, then a planner should not
be forced to specify it. Such an approach permits a least-commitment
approach to planning. Second, in many domains it is simply impossible
to know in planning exactly how long an activity might take. For
example, a common activity of driving a car from one point to another
on a road can take varying amounts of time depending on traffic and
traffic lights. In such circumstances it is more practical to express
upper and lower bounds on durations which naturally lead to intervals
for start and end times. In such domains, flexibility in planning aids
robustness in execution.  Rules

In order for a plan to be valid, it must comply with all rules and
regulations pertinent to the application domain in question. Rules
govern the internal and external relationships of a token. For
example, consider a parameterized predicate describing a transition
from one location to another. Let the parameters be from and to. The
parameters are instantiated on a token as variables whose domain of
values is the set of all locations in a given problem. A rule
governing an internal relationship among token variables might
stipulate that a transition must involve a change in location. This
can be easily expressed as a constraint on the definition of a
predicate of the form: from != to. It is reasonable to further
stipulate that one must be Located somewhere before a transition can
occur, and one must end up Located somewhere when completed. This is
an example of a rule governing an external relationship among
tokens. It specifies a requirement that tokens of the predicate
Located precede and succeed tokens of the predicate Moving.

Figure 2: Internal and external relationships from rules on Moving
Figure 2 illustrates the entities and relations involved in specifying
such a rule on a Moving predicate. The token on which the rule applies
is referred to as the master. Each Located token required by the
master is referred to as a slave. All variables are indicated by name
and their domains are expressed as intervals in the case of temporal
variables and as enumerations for the remainder. Application of a rule
on a token can thus cause slave tokens, variables, and constraints to
be introduced.  Token State Model

The capability of a domain rule to cause a slave token to be created
is a key vehicle through which planning occurs. Semantically, this
rule imposes a requirement for supporting tokens to be in the plan in
order for the master to be valid. There are 2 possibilities to
consider:

The slave is inserted as an active token in the plan. As such, rules
may be activated on the slave, and it may consume resources.

The token is merged with a matching token already in the plan. Once a
slave is merged, the requirement it represents are considered
satisfied. The process of merging passes on all restrictions imposed
on the slave to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible with
the inactive token. For an active token and an inactive token to be
compatible requires that they are instances of the same predicate and
that no intersections between corresponding variables are empty. The
effect of merging is illustrated in Figure 3.

Figure 3: Merging an inactive token on an active token. Domain
restrictions occur in highlighted variables of active token.

On creation of a token, where the commitment has not been made yet to
activate or merge the token, the token is said to be inactive. There
is a nuance to the state model for tokens which relate to the mode of
its creation. If the token is allocated explicitly by an external
actor, rather than internally through rule firing, it may include the
state rejected indicating the planner is permitted to reject the
token. A valid plan can include rejected tokens. This typically arises
where the token represents a goal that is preferable to achieve but
not mandatory. This state is not reachable for slaves since that would
imply selective adherence to the domain model. Figure 4 presents the
state transition diagram for token states relevant for planning. The
transitions are operations on a partial plan which can decide an
outcome for an inactive token. As operations which may arise in
search, they must be reversible during backtracking. The cancellation
operations in each case are also shown.

Figure 4: Token States and Transitions for Planning

The state of a token is embodied with a 5th and final built-in
variable referred to as the state variable. The planner states are
values in the domain of this variable {MERGED, ACTIVE, REJECTED}. The
INACTIVE state is captured by the variable being unbound. If a value
is removed from the domain, that state will not be reachable. For
example, when a slave is created via a rule, the REJECTED value is
removed from its state variable.

\textit{Summary}

This has presented the main elements of the EUROPA planning approach. They are:

@@

\item Temporally scoped predicates and actions (a.k.a. Tokens) to represent
states and actions in time. Note that tokens do not discriminate
between state and action.

\item Token States which are the basis of planning operations on a partial
plan.

\item Constraints to describe relationships among tokens. This provides an
expressive method of describing interactions among tokens in the
context of Temporal Planning.

\item Timelines as a concise abstraction to express the evolution of state
and behavior for a system component. It provides semantics of mutual
exclusion sequencing in time. Other core abstractions are available in
\eu for handling metric resources.

\item Domain rules to describe internal and external relationships on and
between tokens respectively. Rules are applied to active tokens
(master tokens) referred to as masters and typically produce inactive
tokens (slaves) which are a key vehicle through which the planning
process evolves. The rule structure in \eu differs from the more
restrictive commitment to preconditions and effects in classical
planning which prohibits durative actions and disembodied effects
(i.e. effects which may not occur until some temporal distance after
the end of the action).

>>>>

\section{Foundational Concepts}
\label{sec:concepts}

\rx is an adaptive, artificial intelligence based controller and
provides general framework for building reasoning systems for
real-world autonomous vehicles. At MBARI \rx is used for AUV control;
another instantiation of the system is being used for control of a
terrestrial personal robot \cite{pr2, Meeussen:2010dn}. The
development of \rx has been targeted at surveying a number of
oceanographic features which are dynamic and unpredictable
spatio-temporally. Typically this requires our AUVs to balance the
goals of coverage spatially while opportunistically following features
of scientific interest and to do so while being operationally aware of
its own limitations in terms of resources (typically the battery state
of charge) and overall proximity to other observational assets for
obtaining scientific ground-truth. We are therefore interested in
representational frameworks that allow robots to pursue long-term
objectives while choosing short-term gain and can gracefully deal with
exogenous or endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use a
temporal constraint-based planner \eu with a demonstrated legacy of
having flown on NASA space missions \cite{jonsson00,bresina05,
  barreiro09}. Our autonomy architecture brings three key innovations
for AUV adaptation: the use of flexible plan representations,
compositional control with the use of partitioned networks and
off-line learning to inform insitu environmental state estimation. In
this section we detail some of the key concepts relevant to flexible
plan representation. Section \ref{sec:arch} has details the
\rx partitioned architecture and estimation driven vehicle
adaptation.

\paragraph {A brief note on \texttt{EUROPA} history and notation} The
earliest origin of what we will call \eu in this chapter, is from the
\texttt{HSTS} planner \cite{mus94} which was used for a number of
demonstrations for space telescopes as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was the
\texttt{LISP}-based\footnote{This was the first (and only to our
  knowledge) software ever to been written in \texttt{LISP} to be
  flown in space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX}, a C++ version of
the planner was implemented at NASA Ames which went on to be used for
the \texttt{MAPGEN} \cite{bresina05} and continues to be used to date,
in the mission-critical uplink process for the command/control of the
Mars Exploration Rovers (MER) mission. A vastly re-factored, higher
performance open source version \cite{europapso} was named \eu and is
the subject of this chapter. Note however, in using the name \eu, we
refer to the \emph{fundamental ideas as well as the planning paradigm}
even if there exists a specific implementation.

\subsection{The Basics}
\label{sec:basics}

\eu uses a \emph{domain model} written in a declarative language
(NDDL: New Domain Description Language), together with initial
conditions and goals also specified in NDDL, to construct a set of
temporal relations that must be true at the start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, \eu can select a set
of conditions that should be true in the future, where some of these
conditions will correspond to actions the agent must take. The planner
can backtrack and try another path during search if a goal cannot be
reached while being capable of discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a \eu domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

% \begin{figure}
% \centering
% \includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
% \caption{\small Temporal relations defined within the planner are
%   based on \texttt{Allen Algebra} relations shown above.}
% \label{fig:allen-algebra}
% \vskip-0.3cm
% \end{figure}

@@

{\scriptsize
  \begin{quote}
[Planning] \emph{is an abstract, explicit deliberation process that chooses and
organizes actions by anticipating their expected outcomes. This
deliberation aims at achieving as best as possible some prestated
objectives. Automated planning is an area of Artificial Intelligence
(AI) that studies this deliberation process computationally.} -- from
\textbf{Automated Planning Theory and Practice} by Ghallab, Nau and
Traverso \cite{ghallab04} 
\end{quote}
}

\begin{figure} \centering
  \includegraphics[scale=0.35]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\vskip+0.1cm
\end{figure}

@@

For such a robot the milk buying plan at the store might look like
that shown in Fig. \ref{fig:shop:shopping}.

@@

\paragraph \textit{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \cite{Bruckera99}, a project consisting of
a set of activities must be scheduled in a way that satisfies minimum
and/or maximum temporal separation constraints. The activity schedule
must also respect fixed limits on the availability of resources
required to perform each activity. In addition to satisfying temporal
and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \cite{Bruckera99} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. \comment{A figure of a
  constraint network preferably of the shopping example would be
  useful here. JB to generate a napkin fig}

Arc Consistency and other, more stringent consistency definitions can
be used to efficiently prove that a \textsf{CP} problem is
unsatisfiable. However, they generally cannot prove that a \textsf{CP}
problem is satisfiable and provide specific variable assignments that
constitute solutions. For that, search algorithms like global
backtracking \cite{hooker05} or local search are used; these use
consistency as an efficient way to prune the search space
\cite{cp06}. In theory, solving a \textsf{CP} problem is NP-Hard
\cite{ghallab04}, but often very efficient in practice using a number
of consistency and search algorithms that are well understood.

\textsf{CP} is usually implemented as part of a programming language
and constraints are usually represented as objects \cite{puget95}. Any
constraint that the user is able to reason about can be introduced
into the system and, as long as the constraint propagation protocols
specified by the host \textsf{CP} system are enforced, it will be
indistinguishable from any other "primitive'' \textsf{CP} constraint,
such as $\leq$ or $\geq$.

\subsection{Constraint-Based Planning}
\label{sec:europa:cp}

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
% mutual exclusion and concurrency using propositions
% \cite{ghallab04}. While CSP representations have traditionally had an
% edge in formulating and representing planning problems,
% % It is straightforward to represent and reason about all of
% % those elements using a CSP representation, as a result there has been
% % some work on doing automated planning while taking advantage of CSP
% % (TODO: ref). Traditionally, the entire planning problem is translated
% % into a CSP and then solved using traditional CSP methods, this
% leading to a formulation where action choices and relationships are
% represented as variables and/or constraints, there are at least two
% major drawbacks:
mutual exclusion and concurrency using propositions \cite{ghallab04}.
Constraint Programming is a framework that lends itself very well to
represent and reason about all of those elements \cite{ghallab94},
consequently, there has been a good amount of work building automated
planners that take advantage of CP techniques.  The most common
approach in constraint-based planning has been to encode the entire
planning problem as a CP problem, where action choices and
relationships are represented as variables and/or constraints
\cite{do01,vanbeek99,vossen99,wolfman99}.  The advantage of using a
pure CP formulation is that standard CP solvers can be then used to
solve the planning problem, however, this approach also has a couple
of significant drawbacks:

@@

These drawbacks have been addressed by a small number of planning
frameworks like \texttt{Descartes} \cite{joslin95,joslin96},
\texttt{IxTet} \cite{ghallab87,lemai04}, and CAIP (Constraint-Based
Interval and Attribute Planning)\cite{mus94,frank2003}.  These
frameworks take advantage of a CSP representation to deal with time
and resources, at the same time maintaining an explicit representation
of the elements relevant to planning (action choices, causal
relationships, etc). This combination results in a planning approach
that can deal with the type of constraints found in real-life
problems, while allowing for intuitive problem descriptions and a
natural way to express search algorithms and heuristics. \eu is an
implementation of the CAIP framework. In CAIP, planning problems are
specified in terms of:

@@

\item \textbf{Intervals}: entities that have a temporal scope (that
  is, a beginning and an end in time) and a set of
  attributes. Intervals are used to represent actions and states in
  the planning domain.

\item \textbf{Variables and Constraints}: Variables represent Interval
  attributes and Constraints enable a direct representation of
  temporal, resource and any other the restrictions that a valid plan
  must comply with.

\item \textbf{Domain Configuration Rules}: this is an explicit
  representation of the conceptual relationships between actions and
  states in the problem domain, for instance, for a Shopping Agent to
  be able to purchase a product, it must be at a location where the
  product is available.
\end{enumerate}

To generate plans, the CAIP framework uses a plan-space search
approach \cite{ghallab04}, where a partial plan is evolved until all
open goals are achieved, and all inconsistencies are removed, the
specific search algorithm used by \eu will be covered in detail below.

\input{constraints-table}
